---
title: "San Francisco to the Bay"
subtitle: "Measuring the Impact of Transit-Oriented Development"
author: "Oliver Atwood + Dave Drennan + Trevor Kapuvari"
date: '2023-09-21'
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
    
---

# Motivation
Transit is the backbone of development - it influences how and where development occurs, and who lives nearby. But the ways in which transit, demographics, and development are linked varies. This study investigates these links between transit and other factors of the urban landscape through an investigation of four markers: population density, median household income, total population, and median rent.

Our initial work focused on San Francisco county, but given the regional draw of the city and the role that transit plays in linking people from the surrounding area to the city, we expanded the scope to include the entire Bay Area, which is served by the Bay Area Rapid Transit (BART) system. This allowed us to contextualize the unique conditions of San Francisco within the region. In the face of severe income inequality and the housing crisis, it is vital that government officials in the Bay Area region recognize the important role that TOD plays in linking lower-income individuals to employment.

=======

# Setup

```{r setup packages, message = FALSE}
# Clear Environment & Set Chunk Options
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load Libraries
library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)
library(RSocrata)
library(dplyr)
library(stringr)
library(units)
library(gridExtra)
library(autoNumCaptions)

# we don't want scientific notation
options(scipen=999)

# set default class to return the spatial data as sf objects by default
options(tigris_class = "sf")

# load custom functions from the book
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#f0f9e8","#bae4bc","#7bccc4","#43a2ca","#0868ac")

census_api_key("e13d5be5cb48d927009e0dca0af99d21918d514f", overwrite = TRUE)

```

### Wrangling 2009 and 2017 Census Tract ACS data for San Francisco and the Bay Area

```{r tractsData, results = "hide"}
#2009 ACS data for San Francisco only

tracts09_sf <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E",
                        "B19013_001E",
                        "B25058_001E"), 
          year=2009, state=06, county="075", 
          geometry=TRUE, output="wide") %>%
  st_transform('EPSG:2227') %>%
  rename(TotalPop = B25026_001E, 
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E) %>% 
  mutate(area = st_area(geometry)) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(year = "2009", 
         density = drop_units((TotalPop / (area * 0.00000003587 ))))

#2017 ACS data for San Francisco only

tracts17_sf <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E",
                        "B19013_001E",
                        "B25058_001E"), 
          year=2017, state=06, county="075", 
          geometry=TRUE, output="wide") %>%
  st_transform('EPSG:2227') %>%
  rename(TotalPop = B25026_001E, 
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E) %>% 
  mutate(area = st_area(geometry)) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(year = "2017", 
         density = drop_units((TotalPop / (area * 0.00000003587 ))))

#2009 ACS data for Bay Area

tracts09_bay <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E",
                        "B19013_001E",
                        "B25058_001E"), 
          year=2009, state=06, county=c("001", "013", "041", "075", "081", "085", "095"), 
          geometry=TRUE, output="wide") %>%
  st_transform('EPSG:2227') %>%
  rename(TotalPop = B25026_001E, 
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E) %>% 
  mutate(area = st_area(geometry)) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(year = "2009", 
         density = drop_units((TotalPop / (area * 0.00000003587 ))))

#2017 ACS data for Bay Area

tracts17_bay <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E",
                        "B19013_001E",
                        "B25058_001E"), 
          year=2017, state=06, county=c("001", "013", "041", "075", "081", "085", "095"), 
          geometry=TRUE, output="wide") %>%
  st_transform('EPSG:2227') %>%
  rename(TotalPop = B25026_001E, 
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E) %>% 
  mutate(area = st_area(geometry)) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(year = "2017", 
         density = drop_units((TotalPop / (area * 0.00000003587 ))))

# R bind tracts for '09 and '17
allTracts_sf <- rbind(tracts09_sf,tracts17_sf)
allTracts_bay <- rbind(tracts09_bay,tracts17_bay)
```

### Wrangling Transit Open Data

```{r Wrangle Transit Data, results = 'hide'}
BART_Stops_bay <- st_read("https://raw.githubusercontent.com/olivegardener/musa_5080_2023/main/Week_2/data/BART_System.kml")

BART_Stops_bay <- BART_Stops_bay %>%
  st_transform(2227)

sf_outline <- st_union(tracts09_sf)

# Find the points that are within the boundary
indices <- st_within(BART_Stops_bay, sf_outline, sparse = FALSE)

# Create a new sf object containing only the points within the boundary
BART_Stops_sf <- BART_Stops_bay[as.vector(indices), ]

```

### Cropping the extent to show relevant areas

```{r Crop Size}
#Buffer Bart stops for Bounding Box
Buffer4Crop <- st_union(st_buffer(BART_Stops_bay, (15*2640))) %>%
      st_sf() %>%
      mutate(Legend = "Crop Buffer")

# Get the bounding box of Buffer4Crop
bbox_Buffer4Crop <- st_bbox(Buffer4Crop)

# Crop allTracts using the bounding box of Buffer4Crop
cropped_allTracts_bay <- st_crop(allTracts_bay, bbox_Buffer4Crop)


#Crop SF only, no Farallon Islands
cropped_allTracts_sf <- st_crop(allTracts_sf, xmin = 6025077, xmax = 5957122, ymin = 2080091, ymax = 2137110)

```

### Creating buffers

```{r BART Stops Buffer}

BART_Stops_bay_Buffers <- 
  rbind(
    st_buffer(BART_Stops_bay, 2640) %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
    st_union(st_buffer(BART_Stops_bay, 2640)) %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))

BART_Stops_sf_Buffers <- 
  rbind(
    st_buffer(BART_Stops_sf, 2640) %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
    st_union(st_buffer(BART_Stops_sf, 2640)) %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))

buffer_bay <- filter(BART_Stops_bay_Buffers, Legend =="Unioned Buffer")

buffer_sf <- filter(BART_Stops_sf_Buffers, Legend =="Unioned Buffer")

```

### Selecting TOD-designated Census tracts by centroids

```{r Select Centroids}
selectCentroids_bay <-
  st_centroid(tracts09_bay)[buffer_bay,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts09_bay, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Select by Centroids")

selectCentroids_sf <-
  st_centroid(tracts09_sf)[buffer_sf,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts09_sf, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Select by Centroids")

```

### Mapping Census tracts near transit stations

```{r Stations with Tract Selection}
TOD_sf <- ggplot() +
  geom_sf(data=selectCentroids_sf, aes(fill = TotalPop)) +
  geom_sf(data=BART_Stops_sf, show.legend = "point") +
  scale_fill_viridis_c() +
  mapTheme()

TOD_Bay <- ggplot() +
  geom_sf(data=selectCentroids_bay, aes(fill = TotalPop)) +
  geom_sf(data=BART_Stops_bay, show.legend = "point", size = 0.5) +
  scale_fill_viridis_c() +
  mapTheme()

# Display side by side
grid.arrange(TOD_sf, TOD_Bay, ncol=2)

```

# Comparing Indicators

```{r Indicator Maps}
allTracts.group_bay <- 
  rbind(
    st_centroid(allTracts_bay)[buffer_bay,] %>%
      st_drop_geometry() %>%
      left_join(allTracts_bay) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts_bay)[buffer_bay, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts_bay) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
      mutate(MedRent.inf = ifelse(year == "2009", MedRent * 1.15, MedRent)) %>%
      mutate(MedHHInc.inf = ifelse(year == "2009", MedHHInc * 1.15, MedHHInc)) %>%
      st_join(BART_Stops_bay)

allTracts.group_sf <- 
  rbind(
    st_centroid(cropped_allTracts_sf)[buffer_sf,] %>%
      st_drop_geometry() %>%
      left_join(cropped_allTracts_sf) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(cropped_allTracts_sf)[buffer_sf, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(cropped_allTracts_sf) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
      mutate(MedRent.inf = ifelse(year == "2009", MedRent * 1.15, MedRent)) %>%
      mutate(MedHHInc.inf = ifelse(year == "2009", MedHHInc * 1.15, MedHHInc)) %>%
      st_join(BART_Stops_sf)

```

## Time/Space Groups

### TOD vs. Non TOD

In this study, TOD areas are determined by which Census tracts are within half a mile of a BART stop. 

```{r TOD and Non-TOD map sf}
ggplot(allTracts.group_sf)+
    geom_sf(data = st_union(cropped_allTracts_sf))+
    geom_sf(aes(fill = TOD)) +
    geom_sf(data = BART_Stops_sf, col = "black")+
    labs(title = "TOD vs Non-TOD",
         fill = "Type") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))+
    gg_figure_caption(
      caption = "TOD vs Non-TOD tract selection by year"
    )
```

```{r density map sf}
ggplot(allTracts.group_sf)+
    geom_sf(data = st_union(cropped_allTracts_sf))+
    geom_sf(aes(fill = density)) +
    geom_sf(data = st_union(selectCentroids_sf), lwd = 0.5, col = "red",  fill = "NA") +
    geom_sf(data = BART_Stops_sf, col = "black")+
    scale_fill_gradient(low = "white",high = "red4") +
    labs(title = "Population Density",
         subtitle = "BART stops shown as points; TOD tracts selection outlined in red",
         fill = "Per Square Mile") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))+
    gg_figure_caption(
      caption = "Population density per square mile by year"
    )
```
<br>As these maps show, the NIMBY city hates density and has successfully fought it for years.

```{r medhhinc map sf}
ggplot(allTracts.group_sf)+
    geom_sf(data = st_union(cropped_allTracts_sf))+
    geom_sf(aes(fill = MedHHInc)) +
    geom_sf(data = st_union(selectCentroids_sf), lwd = 0.5, col = "red",  fill = "NA") +
    geom_sf(data = BART_Stops_sf, col = "black")+
    scale_fill_gradient(low = "white",high = "darkseagreen4") +
    labs(title = "Median Household Income",
         subtitle = "BART stops shown as points; TOD tracts selection outlined in red \nNote: 2009 dollars inflation adjusted to 2017 dollars",
         fill = "Dollars") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))+
    gg_figure_caption(
      caption = "Median household income by year"
    )
```
<br>This map shows a noticeable increase in mean household income. Interestingly, the areas within the red border (TOD tracts) appear to have increased less than the surrounding area.

```{r total pop map sf}
ggplot(allTracts.group_sf)+
    geom_sf(data = st_union(cropped_allTracts_sf))+
    geom_sf(aes(fill = TotalPop)) +
    geom_sf(data = st_union(selectCentroids_sf), lwd = 0.5, col = "red",  fill = "NA") +
    geom_sf(data = BART_Stops_sf, col = "black")+
    #geom_sf(data = buffer_sf, lwd = 1, col = "darkgoldenrod1",  fill = "NA") +
    scale_fill_gradient(low = "white",high = "darkorchid4") +
    labs(title = "Total Population",
         subtitle = "BART stops shown as points; TOD tracts selection outlined in red",
         fill = "Count") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))+
    gg_figure_caption(
      caption = "Total population by year"
    )
```
<br>This plot shows an increase in overall population, mostly on the east side of the city.

```{r medrent map sf}
ggplot(allTracts.group_sf)+
    geom_sf(data = st_union(cropped_allTracts_sf))+
    geom_sf(aes(fill = MedRent)) +
    geom_sf(data = st_union(selectCentroids_sf), lwd = 0.5, col = "red",  fill = "NA") +
    geom_sf(data = BART_Stops_sf, col = "black")+
    #geom_sf(data = buffer_sf, lwd = 1, col = "darkgoldenrod1",  fill = "NA") +
    scale_fill_gradient(low = "white",high = "cyan4") +
    labs(title = "Median Rent",
         subtitle = "BART stops shown as points; TOD tracts selection outlined in red \nNote: 2009 dollars inflation adjusted to 2017 dollars",
         fill = "Dollars") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))+
    gg_figure_caption(
      caption = "Median rent by year"
    )
```

<br>As with the household income plots, we see an overall increase in Median rent in San Francisco, with the area inside the TOD zone increasing less than the surrounding area.

### Plotting indicators across time and space
```{r TOD Indicator Table bay}
allTracts.Summary_bay <- 
  st_drop_geometry(allTracts.group_bay) %>%
  group_by(year, TOD) %>%
  summarize(Population = mean(TotalPop, na.rm = T),
            Income = mean(MedHHInc, na.rm = T),
            Rent = mean(MedRent, na.rm = T),
            Density = mean(density, na.rm = T))

allTracts.Summary_sf <- 
  st_drop_geometry(allTracts.group_sf) %>%
  group_by(year, TOD) %>%
  summarize(Population = mean(TotalPop, na.rm = T),
            Income = mean(MedHHInc, na.rm = T),
            Rent = mean(MedRent, na.rm = T),
            Density = mean(density, na.rm = T))

```

```{r TOD indicator Plots bay, fig.height = 10}

bay_plot <- allTracts.Summary_bay %>%
  gather(Variable, Value, -year, -TOD) %>%
  ggplot(aes(year, Value, fill = TOD)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, scales = "free", ncol=4) +
  scale_fill_manual(values = c("#bae4bc", "#0868ac")) +
  labs(title = "Bay Area") +
  plotTheme() + theme(legend.position="bottom", plot.margin = unit(c(2,1,1,1),"cm"))+
  xlab("Year")+
  gg_figure_caption(
      caption = "Bay Area TOD analysis indicators"
  )

sf_plot <- allTracts.Summary_sf %>%
  gather(Variable, Value, -year, -TOD) %>%
  ggplot(aes(year, Value, fill = TOD)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, scales = "free", ncol=4) +
  scale_fill_manual(values = c("#bae4bc", "#0868ac")) +
  labs(title = "San Francisco") +
  plotTheme() + theme(legend.position="bottom", plot.margin =unit(c(1,1,2,1),"cm"))+
  xlab("Year")+
  gg_figure_caption(
      caption = "San Francisco TOD analysis indicators"
  )

grid.arrange(sf_plot, bay_plot)
```

## TOD Indicator Tables

### San Francisco

```{r kable table sf}
allTracts.Summary_sf %>%
  unite(year.TOD, year, TOD, sep = ": ", remove = T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Value = round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kable(caption = table.caption("This table compares TOD and non-TOD indicators across years for San Francisco")) %>%
  kable_styling()
```

### Bay Area

```{r kable table bay}
allTracts.Summary_bay %>%
  unite(year.TOD, year, TOD, sep = ": ", remove = T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Value = round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kable(caption = table.caption("This table compares TOD and non-TOD indicators across years for the Bay Area")) %>%
  kable_styling()
```

<br>The observations we have made spatially in the plots seem to bear out in these comparative bar charts and tables. Here we have expanded the study area to include the greater Bay Area, based on the theory that transit has a regional impact on demographics and the built environment. For three of the indicators, the effects are relatively similar, with TOD tracts having higher density and lower income and rent. The effect of transit on population is more varied, with slightly more people living in the TOD zone than in parcels outside the TOD zone for San Francisco, but the opposite story for the wider Bay Area, where more people are living outside the TOD zone than in.

## Graduated Symbol Maps
### San Francisco

```{r graduated symbol map sf}
Stops_Individual_sf <- st_buffer(BART_Stops_sf, 2640)

graduated_stops_sf <-
  st_join(Stops_Individual_sf, st_centroid(cropped_allTracts_bay)) %>%
  group_by(Name) %>%
  summarize(pop_0.5mi_sum = sum(TotalPop, na.rm = TRUE),
            rent_0.5mi_mean = round(mean(MedRent, na.rm = TRUE), digits = 0))

```

```{r graduated map sf setup}
#Calc centroids
tracts_centroids <- st_centroid(cropped_allTracts_sf)

# Initialize an empty vector to store distances
nearest_distances <- vector("numeric", length = nrow(tracts_centroids))

# Loop through each centroid to find the nearest point in FIRST and calculate the distance
for (i in 1:nrow(tracts_centroids)) {
  centroid <- tracts_centroids[i, ]
  nearest_index <- st_nearest_feature(centroid, graduated_stops_sf)
  nearest_point <- graduated_stops_sf[nearest_index, ]
  distance <- st_distance(centroid, nearest_point)
  
  # Store the distance
  nearest_distances[i] <- as.numeric(distance/5280)
}

# Add the distances to the DATABIZNESS dataframe
cropped_allTracts_sf$Distance_to_BART <- nearest_distances
```

```{r graduated income symbol map sf}
# Define a common theme to remove axis text and ticks
common_theme <- theme(
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  axis.title = element_blank(),
  legend.position = "bottom",
  legend.box = "vertical" # Move legend to bottom
)

# Create the first plot
stops_pop_sf <- ggplot() +
  geom_sf(data = cropped_allTracts_sf, 
          aes(fill = Distance_to_BART),  # Fill color based on Dist2Bart
          color = NA,
          alpha = 0.5) +
  geom_sf(data = cropped_allTracts_sf, fill = NA, alpha = 0.5) +
  geom_sf(data = st_centroid(graduated_stops_sf), 
          aes(size = pop_0.5mi_sum),
          color = 'black',
          alpha = 0.75) +
  geom_sf(data = st_centroid(graduated_stops_sf), 
          aes(),
          color = 'black',
          shape = 3,
          size = 0.5) +
  scale_size_continuous(range = c(0.1, 4)) +
  scale_fill_distiller(palette = "Spectral") +
  common_theme +
  labs(title = "Population Within 0.5 miles of BART",  # Add title
       size = "Population", # Legend title
       fill = "Distance to Bart (mi)") +   # legend title
  gg_figure_caption(
      caption = "Proportional population by BART Stop (San Francisco)",
      caption.width = 50
  )

# Create the second plot
stops_rent_sf <- ggplot() +
  geom_sf(data = cropped_allTracts_sf, 
          aes(fill = Distance_to_BART),  # Fill color based on Dist2Bart
          color = NA,
          alpha = 0.5) +
  geom_sf(data = cropped_allTracts_sf, fill = NA, alpha = 0.5) +
  geom_sf(data = st_centroid(graduated_stops_sf), 
          aes(size = rent_0.5mi_mean),
          color = 'black',
          alpha = 0.75) +
  geom_sf(data = st_centroid(graduated_stops_sf), 
          aes(),
          color = 'black',
          shape = 3,
          size = 0.5) +
  scale_size_continuous(range = c(0.1, 4)) +
  scale_fill_distiller(palette = "Spectral") +  
  common_theme +
  labs(title = "Mean Rent Within 0.5 Miles of BART",  # Add title
       size = "Mean Rent",  # Legend title
       fill = "Distance to Bart (mi)") +  # Legend title
  gg_figure_caption(
      caption = "Proportional mean rent by BART Stop (San Francisco)",
      caption.width = 50
  )
  
# Arrange the plots side by side
grid.arrange(stops_pop_sf, stops_rent_sf, ncol = 2)

```  
<br>Here we can see the spatial variation of population and median rent in SF for each BART stop. Of interest here is the effect of Center City on population and rent. Interestingly, for the BART stops nearest to Center City, the population of TOD census tracts is relatively higher and the mean rent is relatively lower.

### Bay Area

```{r graduated symbol map bay}
Stops_Individual_bay <- st_buffer(BART_Stops_bay, 2640)

graduated_stops_bay <-
  st_join(Stops_Individual_bay, st_centroid(cropped_allTracts_bay)) %>%
  group_by(Name) %>%
  summarize(pop_0.5mi_sum = sum(TotalPop, na.rm = TRUE),
            rent_0.5mi_mean = round(mean(MedRent, na.rm = TRUE), digits = 0))

```

```{r graduated map bay setup, message = FALSE}
#Calc centroids
tracts_centroids <- st_centroid(cropped_allTracts_bay)

# Initialize an empty vector to store distances
nearest_distances <- vector("numeric", length = nrow(tracts_centroids))

# Loop through each centroid to find the nearest point in FIRST and calculate the distance
for (i in 1:nrow(tracts_centroids)) {
  centroid <- tracts_centroids[i, ]
  nearest_index <- st_nearest_feature(centroid, graduated_stops_bay)
  nearest_point <- graduated_stops_bay[nearest_index, ]
  distance <- st_distance(centroid, nearest_point)
  
  # Store the distance
  nearest_distances[i] <- as.numeric(distance/5280)
}

# Add the distances to the DATABIZNESS dataframe
cropped_allTracts_bay$Distance_to_BART <- nearest_distances
```

```{r graduated income symbol map bay, fig.width=12, fig.height=6.75}
# Define a common theme to remove axis text and ticks
common_theme <- theme(
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  axis.title = element_blank(),
  legend.position = "bottom",
  legend.box = "vertical" # Move legend to bottom
)

# Create the first plot
stops_pop_bay <- ggplot() +
  geom_sf(data = cropped_allTracts_bay, 
          aes(fill = Distance_to_BART),  # Fill color based on Dist2Bart
          color = NA,
          alpha = 0.5) +
  geom_sf(data = st_centroid(graduated_stops_bay), 
          aes(size = pop_0.5mi_sum),
          color = 'black',
          alpha = 0.5) +
  geom_sf(data = st_centroid(graduated_stops_bay), 
          aes(),
          color = 'black',
          shape = 3,
          size = 0.25) +
  scale_size_continuous(range = c(0.1, 4)) +
  scale_fill_distiller(palette = "Spectral") +
  common_theme +
  labs(title = "Population Within 0.5 miles of BART",  # Add title
       size = "Population",
       fill = "Distance to Bart (mi)")+   # legend title
  gg_figure_caption(
      caption = "Proportional population by BART Stop (Bay Area)",
      caption.width = 50
  )

# Create the second plot
stops_rent_bay <- ggplot() +
  geom_sf(data = cropped_allTracts_bay, 
          aes(fill = Distance_to_BART),  # Fill color based on Dist2Bart
          color = NA,
          alpha = 0.5) +
  geom_sf(data = st_centroid(graduated_stops_bay), 
          aes(size = rent_0.5mi_mean),
          color = 'black',
          alpha = 0.5) +
  geom_sf(data = st_centroid(graduated_stops_bay), 
          aes(),
          color = 'black',
          shape = 3,
          size = 0.25) +
  scale_size_continuous(range = c(0.1, 4)) +
  scale_fill_distiller(palette = "Spectral") +  
  common_theme +
  labs(title = "Mean Rent Within 0.5 Miles of BART",  # Add title
       size = "Mean Rent",
       fill = "Distance to Bart (mi)") +  # Legend title
  gg_figure_caption(
      caption = "Proportional mean rent by BART Stop (Bay Area)",
      caption.width = 50
  ) 

# Arrange the plots side by side
grid.arrange(stops_pop_bay, stops_rent_bay, ncol = 2)

```  
<br>Here are the same plots for the Bay Area. These show some areas of the east bay where population and rent in the TOD zone are relatively low. These could be good areas for the development of affordable housing.

## Multiple Ring Buffer Analysis

```{r multipleRingBuffer bay}

BART_MRB_sf <- multipleRingBuffer(st_union(BART_Stops_sf), 47520, 2640)

allTracts.rings_sf <-
  st_join(st_centroid(dplyr::select(allTracts_sf, GEOID, year)),
          BART_MRB_sf) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts_sf, GEOID, MedRent, year), 
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance = distance / 5280) #convert to miles

allTracts.rings.summary_sf <- st_drop_geometry(allTracts.rings_sf) %>%
    group_by(distance, year) %>%
    summarize(Mean_Rent = mean(MedRent, na.rm=T))

sf_lines <- ggplot(allTracts.rings.summary_sf,
       aes(distance, Mean_Rent, colour=year)) +
      geom_point(size=3) + 
  geom_line(size=2)+
  scale_x_continuous(limits = c(0, 10))+
  scale_y_continuous(limits = c(0, 2100))+
  xlab("Distance (mi)")+
  ylab("Mean Rent ($)")+
  gg_figure_caption(
      caption = "Mean Rent as a function of distance to BART stop (San Francisco)") +
  theme(plot.caption = element_text(hjust = 0))

BART_MRB_bay <- multipleRingBuffer(st_union(BART_Stops_bay), 47520, 2640)

allTracts.rings_bay <-
  st_join(st_centroid(dplyr::select(allTracts_bay, GEOID, year)),
          BART_MRB_bay) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts_bay, GEOID, MedRent, year), 
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance = distance / 5280) #convert to miles

allTracts.rings.summary_bay <- st_drop_geometry(allTracts.rings_bay) %>%
    group_by(distance, year) %>%
    summarize(Mean_Rent = mean(MedRent, na.rm=T))

bay_lines <- ggplot(allTracts.rings.summary_bay,
       aes(distance, Mean_Rent, colour=year)) +
      geom_point(size=3) + 
  geom_line(size=2)+
  scale_x_continuous(limits = c(0, 10))+
  scale_y_continuous(limits = c(0, 2100))+
  xlab("Distance (mi)")+
  ylab("Mean Rent ($)")+
  gg_figure_caption(caption = "Mean Rent as a function of distance to BART stop (Bay Area)") +
  theme(plot.caption = element_text(hjust = 0))

grid.arrange(sf_lines, bay_lines)
```
<br>These charts show that proximity to BART has a slight *negative* influence on mean rent. In San Francisco this influence is more variable whereas the Bay Area has a more linear increase in rent with distance from BART stations.


# Conclusion

Ultimately, our analysis does not indicate that residents are willing to pay more to live in TOD areas versus non-TOD areas. The presence of public transit in both San Francisco and the Bay Area appears to be correlated with areas of higher density and lower rent.

In a region facing both severe income inequality and a devastating housing crisis, density and affordability are of paramount importance. The findings of this research suggest that TOD is not an especially profitable pursuit for developers, underscoring the importance the public sector's role in incentivizing TOD through tax breaks and direct subsidy. Policy-makers should seek out areas in the region to incentivize the construction of affordable housing near transit to help address the housing crisis.
