---
title: "Lab 2: Why Start With Indicators?"
author: "Matt Harris & Michael Fichman"
date: '2022-08-22'
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup clear environment, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

### Load packages and functions

```{r setup packages, warning = FALSE}
# Load Libraries

library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)
library(RSocrata)
library(dplyr)
library(stringr)

# we don't want scientific notation
options(scipen=999)

# set default class to return the spatial data as sf objects by default
options(tigris_class = "sf")

# load custom functions from the book
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#f0f9e8","#bae4bc","#7bccc4","#43a2ca","#0868ac")

census_api_key("e13d5be5cb48d927009e0dca0af99d21918d514f", overwrite = TRUE)

```


```{r tracts09}
tracts09 <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E","B25058_001E",
                        "B06012_002E"), 
          year=2009, state=06, county=075, 
          geometry=TRUE, output="wide") %>%
  st_transform('EPSG:2227') %>%
  rename(TotalPop = B25026_001E, 
         Whites = B02001_002E,
         FemaleBachelors = B15001_050E, 
         MaleBachelors = B15001_009E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E) %>% 
  mutate(area = st_area(geometry)) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop),0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2009", density = (TotalPop / (area * 0.00000003587 ))) %>%
  dplyr::select(-Whites, -FemaleBachelors, -MaleBachelors, -TotalPoverty) 

```

```{r tracts17 results='hide'}
tracts17 <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E","B25058_001E",
                        "B06012_002E"), 
          year=2017, state=06, county=075, 
          geometry=TRUE, output="wide") %>%
  st_transform('EPSG:2227') %>%
  rename(TotalPop = B25026_001E, 
         Whites = B02001_002E,
         FemaleBachelors = B15001_050E, 
         MaleBachelors = B15001_009E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E) %>% 
  mutate(area = st_area(geometry)) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop),0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2017", density = (TotalPop / (area * 0.00000003587 ))) %>%
  dplyr::select(-Whites, -FemaleBachelors, -MaleBachelors, -TotalPoverty) 
```

<!-- Filter out Farralon Islands for 2017 -->
<!-- ```{r Filter Farralon 2} -->
<!-- # Calculate the centroids -->
<!-- centroids <- st_centroid(tracts17) -->

<!-- # Extract the centroid coordinates -->
<!-- coords <- st_coordinates(centroids) -->

<!-- # Add the coordinates as new columns -->
<!-- tracts17$lon <- coords[, "X"] -->
<!-- tracts17$lat <- coords[, "Y"] -->

<!-- min_lon <- min(tracts17$lon) -->

<!-- tracts17 <- tracts17 %>% -->
<!--    filter(lon > min_lon) -->

<!-- tracts17 <- tracts17 %>% -->
<!--   select(-lat, -lon) -->

<!-- ``` -->

```{r bind tracts}

allTracts <- rbind(tracts09,tracts17)

#Crop Farrallon Island instead?
allTracts <- st_crop(allTracts, xmin = 6025077, xmax = 5957122, ymin = 2080091, ymax = 2137110)
```

### Wrangling Transit Open Data

```{r}
BART_Stops <- st_read("D:/Penn/Comms/musa_5080_2023/Week_2/data/BART_Sta_13.shp")

BART_Stops <- BART_Stops %>%
  st_transform(2227)

sf_outline <- st_union(tracts09)

# Find the points that are within the boundary
indices <- st_within(BART_Stops, sf_outline, sparse = FALSE)

# Create a new sf object containing only the points within the boundary
BART_Stops_SF <- BART_Stops[as.vector(indices), ]

# MUNI_Stops <- read.socrata(
#   "https://data.sfgov.org/resource/i28k-bkz6.json",
#   app_token = "fpOaceiSu5zHVOar6s6P4xNah",
#   email     = "oatwood@upenn.edu",
#   password  = "viqZi4-jucpuh-zimhas"
# )
# 
# MUNI_Stops_sf <- st_as_sf(MUNI_Stops, coords = c("longitude", "latitude"), crs = 4326)
# 
# MUNI_Stops_sf <- MUNI_Stops_sf %>%
#   st_transform(2227)
```

```{r }
BART_Stops_Buffers <- 
  rbind(
    st_buffer(BART_Stops_SF, 2640) %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
    st_union(st_buffer(BART_Stops_SF, 2640)) %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))
```

Let's examine both buffers by making a "[small
multiples](https://www.juiceanalytics.com/writing/better-know-visualization-small-multiples)"
plot. This type of plot shows multiple views of the same data or
geographic area. We accomplish this by using the `facet_wrap()`
{ggplot2} function on the *Legend* column.

```{r}
ggplot() +
  geom_sf(data=allTracts) +
  facet_wrap(pctWhite~year) + 
  mapTheme()
```

```{r}
buffer <- filter(BART_Stops_Buffers, Legend =="Unioned Buffer")
```

```{r}
selectCentroids <-
  st_centroid(tracts09)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts09, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Select by Centroids")
```


Plotting the results of each method

```{r}
ggplot() +
  geom_sf(data=selectCentroids, aes(fill = TotalPop)) +
  geom_sf(data=BART_Stops_SF, show.legend = "point") +
  scale_fill_viridis_c() +
  facet_wrap(~Selection_Type) + 
  mapTheme()
```
### Indicator Maps

We do our centroid joins as above, and then do a "disjoin" to get the
ones that *don't* join, and add them all together. Do this operation and
then examine it. What represents the joins/doesn't join dichotomy? Note
that this contains a correct 2016-2020 inflation calculation

```{r}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
    mutate(MedRent.inf = ifelse(year == "2009", MedRent * 1.14, MedRent)) %>%
      st_join(BART_Stops)

```

### TOD Indicator Tables

```{r}
allTracts.Summary <- 
  st_drop_geometry(allTracts.group) %>%
  group_by(year, TOD) %>%
  summarize(Rent = mean(MedRent, na.rm = T),
            Population = mean(TotalPop, na.rm = T),
            #Percent_White = mean(pctWhite, na.rm = T),
            Density = mean(density, na.rm = T),
            Percent_Poverty = mean(pctPoverty, na.rm = T))

kable(allTracts.Summary) %>%
  kable_styling() %>%
  footnote(general_title = "\n",
           general = "Table 2.2")
```

Let's make some comparisons and speculate about the willingness to pay
and demographics in these areas 2009-2020 (see the 2000 data in the text
too)

Notice how we pipe the kable() command here

```{r}
allTracts.Summary %>%
  unite(year.TOD, year, TOD, sep = ": ", remove = T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Value = round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kable() %>%
  kable_styling() %>%
  footnote(general_title = "\n",
           general = "Table 2.3")
```

### TOD Indicator Plots

Let's create small multiple plots We use the "gather" command (look this
one up please) To go from wide to long Why do we do this?? Notice we can
"pipe" a ggplot call right into this operation!

```{r}
allTracts.Summary %>%
  gather(Variable, Value, -year, -TOD) %>%
  ggplot(aes(year, Value, fill = TOD)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, scales = "free", ncol=5) +
  scale_fill_manual(values = c("#bae4bc", "#0868ac")) +
  labs(title = "Indicator differences across time and space") +
  plotTheme() + theme(legend.position="bottom")
```

#### Examining three submarkets




```{r}
centerCity <-
  st_intersection(
    st_buffer(filter(septaStops, Line == "El"), 2640) %>% st_union(),
    st_buffer(filter(septaStops, Line == "Broad_St"), 2640) %>% st_union()) %>%
  st_sf() %>%
  mutate(Submarket = "Center City")

el <-
  st_buffer(filter(septaStops, Line == "El"), 2640) %>% st_union() %>%
  st_sf() %>%
  st_difference(centerCity) %>%
  mutate(Submarket = "El")

broad.st <-
  st_buffer(filter(septaStops, Line == "Broad_St"), 2640) %>% st_union() %>%
  st_sf() %>%
  st_difference(centerCity) %>%
  mutate(Submarket = "Broad Street")

threeMarkets <- rbind(el, broad.st, centerCity)
```

You can then bind these buffers to tracts and map them or make small
multiple plots

```{r}
allTracts.threeMarkets <-
  st_join(st_centroid(allTracts), threeMarkets) %>%
  st_drop_geometry() %>%
  left_join(allTracts) %>%
  mutate(Submarket = replace_na(Submarket, "Non-TOD")) %>%
  st_sf() 
```


```{r graduated symbol map }
Stops_Individual <- st_buffer(BART_Stops_SF, 2640)
#Crop Farrallon Island ?
tracts17_crop <- st_crop(tracts17, xmin = 6025077, xmax = 5957122, ymin = 2080091, ymax = 2137110)


propmap <-
  st_join(Stops_Individual, st_centroid(tracts17)) %>%
  group_by(STATION) %>%
  summarize(pop_0.5mi_sum = sum(TotalPop),
            rent_0.5mi_mean = round(mean(MedRent), digits = 0))

#propmap <- gather(propmap, key = "Stat", value, 2:3) 
  
ggplot()+
  geom_sf(data = tracts17_crop, fill = "white") +
  geom_sf(data = st_centroid(propmap),
          aes(size = pop_0.5mi_sum)) +
  scale_size_continuous(range = c(1,9))

ggplot()+
  geom_sf(data = tracts17_crop, fill = "white") +
  geom_sf(data = st_centroid(propmap),
          aes(size = rent_0.5mi_mean)) +
  scale_size_continuous(range = c(1,9))


  
```  
  
### Example of using `multipleRingBuffer()` function

```{r multipleRingBuffer}
BART_MRB <- multipleRingBuffer(st_union(BART_Stops_SF), 47520, 2640)

 # ggplot() +
 #    geom_sf(data=BART_MRB) +
 #    geom_sf(data=BART_Stops_SF, size=1) +
 #    geom_sf(data=st_union(tracts17_crop), fill=NA, size=1.2) +
 #    labs(title="Half mile buffers") +
 #    mapTheme()

allTracts.rings <-
  st_join(st_centroid(dplyr::select(allTracts, GEOID, year)),
          BART_MRB) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year), 
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance = distance / 5280) #convert to miles

allTracts.rings.summary <- st_drop_geometry(allTracts.rings) %>%
    group_by(distance, year) %>%
    summarize(Mean_Rent = mean(MedRent, na.rm=T))

ggplot(allTracts.rings.summary,
       aes(distance, Mean_Rent, colour=year)) +
      geom_point(size=3) + 
  geom_line(size=2)



```

```{r}
allTracts <- rbind(tracts20,tracts09)
septa_MRB <- multipleRingBuffer(st_union(septaStops), 47520, 2640)

allTracts.rings <-
  st_join(st_centroid(dplyr::select(allTracts, GEOID, year)),
          septa_MRB) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year), 
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance = distance / 5280) #convert to miles



  ggplot() +
    geom_sf(data=septa_MRB) +
    geom_sf(data=septaStops, size=1) +
    geom_sf(data=st_union(tracts20), fill=NA, size=1.2) +
    labs(title="Half mile buffers") +
    mapTheme()

```

