---
title: "OSM Data Wrangling"
output: html_document
date: "2023-09-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

## R Markdown

```{r}
library(osmdata)
library(sf)
library(ggplot2)
library(raster)
library(dplyr)
library(terra)
library(tidyr)
library(tigris)
library(tidycensus)

# Capture the start time
start_time <- Sys.time()
```

## Setting Parameters
```{r}
# coordinate_system <- 32756
# 
# resolution <- 300

coordinate_system <- 4326
# Target resolution in m
resolution <- 100
# Conversion to degrees
resolution <- (resolution/30)/3600

```


```{r install_census_API_key, warning = FALSE, include=FALSE, eval = TRUE}
# Install Census API Key
tidycensus::census_api_key("a3ebdf1648b7fb21df55df7246d9642f040c0ee0", overwrite = TRUE)
```

```{r}
# Optional: Set options to use sf class
options(tigris_class = "sf")

# Get place data for Illinois
illinois_places <- places(state = "17")

# Filter for Chicago
chicago_boundary <- illinois_places[illinois_places$NAME == "Chicago", ]

# Set the bounding box
bbox <- st_bbox(chicago_boundary)

```

## 2.3. Import Census Info

```{r get_census, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide'}
chicagoCensus <- 
  get_acs(geography = "tract", 
          variables = c("B19013_001", "B25002_003", "B25001_001"), 
          year = 2020, 
          state = "IL", 
          geometry = TRUE, 
          county=c("Cook"),
          output = "wide") %>%
  rename(Med_Inc = B19013_001E,
         Total_Vacancy = B25002_003E,
         Total_Units = B25001_001E) %>%
  mutate(Percent_Vacancy = Total_Vacancy / Total_Units) %>% 
  dplyr::select(Med_Inc, Percent_Vacancy, GEOID, geometry)

```



```{r add_census_tracts , message = FALSE, warning = FALSE}
dat_census <- st_join(dat2 %>%
          filter(is.na(from_longitude) == FALSE &
                   is.na(from_latitude) == FALSE &
                   is.na(to_latitude) == FALSE &
                   is.na(to_longitude) == FALSE) %>%
          st_as_sf(., coords = c("from_longitude", "from_latitude"), crs = 4326),
        chicagoTracts %>%
          st_transform(crs=4326),
        join=st_intersects,
              left = TRUE) %>%
  rename(Origin.Tract = GEOID) %>%
  mutate(from_longitude = unlist(map(geometry, 1)),
         from_latitude = unlist(map(geometry, 2)))%>%
  as.data.frame() %>%
  dplyr::select(-geometry)%>%
  st_as_sf(., coords = c("to_longitude", "to_latitude"), crs = 4326) %>%
  st_join(., chicagoTracts %>%
            st_transform(crs=4326),
          join=st_intersects,
          left = TRUE) %>%
  rename(Destination.Tract = GEOID)  %>%
  mutate(to_longitude = unlist(map(geometry, 1)),
         to_latitude = unlist(map(geometry, 2)))%>%
  as.data.frame() %>%
  dplyr::select(-geometry)
```

## Download OSM Data

In OpenStreetMap (OSM), the highway key is used to tag various types of roads, paths, and other routes. The value for the highway key can vary depending on the type of road or path you are interested in. Here are some common values used with the highway key:

motorway: Major freeways or interstates designed for high-speed vehicular traffic.
trunk: Important roads that aren't motorways. Typically these roads will have fewer points of access, be a higher speed, and a higher capacity than other roads.
primary: Major roads in a country's road network, typically linking large towns and significant suburbs of large cities.
secondary: Roads that link smaller towns and larger villages.
tertiary: Roads that link smaller towns and villages.
unclassified: The OSM default for minor roads.
residential: Roads within a residential area.
service: Roads that provide access to a building, service facility, or parking area.
motorway_link, trunk_link primary_link, secondary_link, tertiary_link: Link roads typically used to connect roads of the named classification with other roads of equal or higher classification, e.g., a slip road connecting a motorway and a trunk road.
living_street: Roads within a residential area where pedestrians have legal priority over cars, speeds are very low, and children can play on the street.
pedestrian: Roads primarily for pedestrians in shopping or residential areas.
track: Roads for agricultural use, gravel roads in the forest, etc. These are often rough and may not be paved.
busway: Roads designated for buses only.
footway: Paths for walking or hiking.
cycleway: Paths designated for cycling.
bridleway: Paths designated for horse riding.

This list is not exhaustive, as there are more specific and less commonly used values, but these cover most general road types. For a complete and up-to-date list, you can refer to the OpenStreetMap Wiki page on highway tags.

```{r Download OSM Data, echo=FALSE}
# Query Roads Data from OSM
roads_data_1 <- opq(bbox) %>%
  add_osm_feature(key = "highway", value = "trunk") %>%
  osmdata_sf()

roads_data_2 <- opq(bbox) %>%
  add_osm_feature(key = "highway", value = "trunk_link") %>%
  osmdata_sf()

# roads_data_3 <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "primary") %>%
#   osmdata_sf()
# 
# roads_data_4 <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "primary_link") %>%
#   osmdata_sf()
# 
# roads_data_5 <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "secondary") %>%
#   osmdata_sf()
# 
# roads_data_6 <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "secondary_link") %>%
#   osmdata_sf()
```

```{r pre-process, echo=FALSE}
# Convert the roads data to a simple feature object
roads_1 <- roads_data_1$osm_lines
roads_utm_1 <- st_transform(roads_1, coordinate_system)
roads_utm_1 <- st_simplify(roads_utm_1, dTolerance = resolution)

roads_2 <- roads_data_2$osm_lines
roads_utm_2 <- st_transform(roads_2, coordinate_system)
roads_utm_2 <- st_simplify(roads_utm_2, dTolerance = resolution)

# roads_3 <- roads_data_3$osm_lines
# roads_utm_3 <- st_transform(roads_3, coordinate_system)
# roads_utm_3 <- st_simplify(roads_utm_3, dTolerance = resolution)
# 
# roads_4 <- roads_data_4$osm_lines
# roads_utm_4 <- st_transform(roads_4, coordinate_system)
# roads_utm_4 <- st_simplify(roads_utm_4, dTolerance = resolution)
# 
# roads_5 <- roads_data_5$osm_lines
# roads_utm_5 <- st_transform(roads_5, coordinate_system)
# roads_utm_5 <- st_simplify(roads_utm_5, dTolerance = resolution)
# 
# roads_5 <- roads_data_5$osm_lines
# roads_utm_5 <- st_transform(roads_5, coordinate_system)
# roads_utm_5 <- st_simplify(roads_utm_5, dTolerance = resolution)

# # Combine roads
# all_roads <- bind_rows(roads_utm_1, roads_utm_2, roads_utm_3, roads_utm_4, roads_utm_5)
# # Keeping only osm_id and geometry columns
# all_roads <- all_roads[, c("osm_id", "geometry")]

# Isolate highways
highway <- bind_rows(roads_utm_1, roads_utm_2)
# Keeping only osm_id and geometry columns
highway <- highway[, c("osm_id", "geometry")]

# Plot using ggplot2
ggplot(data = highway) +
  geom_sf() +
  theme_minimal() +
  labs(title = "Imported Roads")


```

## Rasterize Roads Data
```{r RASTERIZE, fig.width=max}
# Extract extent and crs from cropbox
extent <- ext(cropbox)
crs <- crs(cropbox)

# Create an empty raster based on that extent and CRS
r <- rast(extent, res = resolution, crs = crs)

### ALL ROADS 
# Convert the DataFrame to a SpatVector
all_roads_vect <- vect(all_roads)
# Rasterize the Roads
roads_raster <- rasterize(all_roads_vect, r, fun = "sum")
# Create a Binary Raster
binary_roads_raster <- ifel(roads_raster > 0, 1, 0)
# Calculate Proximity
road_proximity <- distance(binary_roads_raster)

### HIGHWAYS
highway_vect <- vect(highway)
highway_raster <- rasterize(highway_vect, r, fun = "sum")
binary_highway_raster <- ifel(highway_raster > 0, 1, 0)
highway_proximity <- distance(binary_highway_raster)

# Plot the rasters
plot(road_proximity, main = "Proximity to Roads")
plot(highway_proximity, main = "Proximity to highways")

# write rasters
# MAC
# writeRaster(road_proximity, filename = "/Users/oliveratwood/Documents/GitHub/HotspotStoplight/UrbanExpansion/data/roads_proximity.tif", format = "GTiff", overwrite = TRUE)
# writeRaster(highway_proximity, filename = "/Users/oliveratwood/Documents/GitHub/HotspotStoplight/UrbanExpansion/data/highway_proximity.tif", format = "GTiff", overwrite = TRUE)

# WINDOWS
# writeRaster(road_proximity, filename = "C:\\Users\\olive\\Documents\\GitHub\\HotspotStoplight\\UrbanExpansion\\data\\roads_proximity.tif", format = "GTiff", overwrite = TRUE)
# writeRaster(highway_proximity, filename = "C:\\Users\\olive\\Documents\\GitHub\\HotspotStoplight\\UrbanExpansion\\data\\highway_proximity.tif", format = "GTiff", overwrite = TRUE)

```

```{r timer_end}
# Capture the end time
end_time <- Sys.time()

# Calculate and print the runtime
runtime <- (end_time - start_time)/60
print(paste("Total runtime (mins):", runtime))
```

