---
title: "Cut Code"
author: "Dave Drennan"
date: "2023-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
##OA Local files
# TreeCanopy <- st_read("/Users/oliveratwood/Box Sync/PPA Midterm/data/TreeCanopy.geojson")
# landuse <- st_read("/Users/oliveratwood/Box Sync/PPA Midterm/data/Land_Use.geojson")%>%
#   st_transform(crs = coordinate_system)

catchments <- st_read("data/School_Catchment/Catchment_ES_2021-22") %>% st_transform(crs = coordinate_system)
Redlined <- st_read("data/redlining.geojson")
Vacant_Lots <- st_read("data/Vacant_Lots.geojson")
Recycling_Diversion_Rate <- st_read("data/Recycling_Diversion_Rate.geojson")
Hospitals <- st_read("data/Hospitals.geojson")
Regional_Rail_Stations <- st_read("data/Regional_Rail_Stations.geojson")
Existing_Trails <- st_read("data/Existing_Trails.geojson")
Traffic_Collisions <- st_read("data/PhillyHealth_Collisions.geojson")
trees <- st_read("data/trees.geojson") %>% mutate(TREE_DBH = as.factor(TREE_DBH))
bike_network <- st_read("data/bike_network.geojson")
parks <- st_read("data/parks.geojson")
uni <- st_read("data/uni.geojson")
neighborhoods <- st_read("data/neighborhoods.geojson")
landcare_lots <- st_read("data/landcare_lots.geojson")
tracts2020 <- st_read("data/tracts2020.geojson")
septa <- st_read("data/septa.geojson")
planning_districts <- st_read("data/Planning_Districts.geojson") %>% st_transform(coordinate_system)
septa_bus <- septa %>% 
  filter(mode == "Bus")
septa_trolley <- septa %>% 
  filter(mode == "Trolley")
septa_highspeed <- septa %>% 
  filter(mode == "Highspeed")

```


### Compute percent canopy cover within 500ft
```{r warning = FALSE}
# # Create buffers around each house
# HouseBuffers <- st_buffer(Houses, dist = (0.25*5280))
# 
# # Initialize an empty vector to store the canopy coverage percent for each house
# CanopyPct100ft <- vector("numeric", length = nrow(Houses))
# 
# # Loop through each house buffer
# for (i in 1:nrow(HouseBuffers)) {
# # Calculate the intersection area between the current house buffer and the tree canopy
# intersection_area <- sum(st_area(st_intersection(HouseBuffers[i, ], TreeCanopy)))
#   
# # Calculate the canopy coverage percent for the current house buffer
# CanopyPct100ft[i] <- (intersection_area / st_area(HouseBuffers[i, ])) * 100}
# 
# # Add the canopy coverage percent values to the Houses data frame
# Houses$CanopyPctQuMi <- CanopyPct100ft

```


### Compute min dist to trails and bike network
```{r}
# these are proxy values to speed runs, replace with commented section in final version
Houses$dist2trail <- 5
Houses$dist2bike <- 5
Houses$dist2park <- 5

# ## Trails
# # Compute distances
# distances <- st_distance(Houses, Existing_Trails)
# # Compute the minimum distance for each point to the nearest line
# min_distances <- apply(distances, 1, min)
# # Join the minimum distances to the attribute table of the points
# Houses$dist2trail <- min_distances
# 
# ## Bike Network
# distances <- st_distance(Houses, bike_network)
# min_distances <- apply(distances, 1, min)
# Houses$dist2bike <- min_distances
# 
# # Parks
# distances <- st_distance(Houses, parks)
# min_distances <- apply(distances, 1, min)
# Houses$dist2park <- min_distances

```

Let's try making a model with all of our variables, to start.

```{r model 1}
Houses_Clean <- dplyr::select(Houses, -musaID)

reg1 <- lm(sale_price ~ ., data = select_if(st_drop_geometry(Houses_Clean), is.numeric))
summary(reg1)

```
Let's try narrowing down the variables based on a correlogram

From this correlogram, we can see that there are some potentially co-linear variables and variables with low correlation with sale_price. Let's remove these variables and make a model with this smaller set.

```{r model 2}
Houses2 <- Houses_Clean %>% dplyr::select(-number_stories, -age, -PopDensity_ACS, -Car2WorkPct_ACS, -dist2trail, -dist2park, -hospitals_nn1, -rr_station_nn1, -septa_trolley_nn1, -septa_highspeed_nn1, -unis_nn1, -landcare_lots_nn5)

reg2 <- lm(sale_price ~ ., data = select_if(st_drop_geometry(Houses2), is.numeric))
summary(reg2)


coefficients(reg2)
```
Our R-squared variable only decreased by 0.01. Not bad at all, given the number of variables removed. There are a lot of variables in this model that appear to have low significance, let's remove them.


```{r model 3}
Houses3 <- Houses2 %>% dplyr::select(-garage_spaces, -MedHHInc_ACS, -MedRent_ACS, -PctWhite_ACS, -dist2bike, -collisions_nn3, -vacant_lots_nn5)

reg3 <- lm(sale_price ~ ., data = select_if(st_drop_geometry(Houses3), is.numeric))
summary(reg3)

```

```{r model 4}
Houses4 <- Houses3 %>% dplyr::select(-RecyclingRate)
```


We weren't taking into account spatial adjustments - I think the whittling down threw off the original regressions. Following chunk creates a dataset of only variables we care about and includes redlining - was having trouble with the factor level with neighborhoods and catchments.
```{r regression final}

Houses_Clean <- Houses_Clean %>%
  mutate(
    PctBach_ACS = PctBach_ACS * 100,
    exterior_condition = ifelse(is.na(exterior_condition), median(exterior_condition, na.rm = TRUE), exterior_condition),
    number_of_bathrooms = ifelse(is.na(number_of_bathrooms), median(number_of_bathrooms, na.rm = TRUE), number_of_bathrooms),
    number_of_bedrooms = ifelse(is.na(number_of_bedrooms), median(number_of_bedrooms, na.rm = TRUE), number_of_bedrooms),
    interior_condition = ifelse(is.na(interior_condition), median(interior_condition, na.rm = TRUE), interior_condition),
    fireplaces = ifelse(is.na(fireplaces), median(fireplaces, na.rm = TRUE), fireplaces),
    total_area = ifelse(is.na(total_area), median(total_area, na.rm = TRUE), total_area)
  )


regression_data <- Houses_Clean %>% dplyr::select(
                                                  sale_price,
                                                  exterior_condition, 
                                                  fireplaces, 
                                                  interior_condition, 
                                                  number_of_bathrooms, 
                                                  number_of_bedrooms, 
                                                  #number_of_rooms, 
                                                  total_area, 
                                                  total_livable_area, 
                                                  PctBach_ACS, 
                                                  septa_bus_nn5, 
                                                  #Neighborhood, 
                                                  #holc_grade,
                                                  dist
                                                  #ES_Name
                                                  ) 

regression <- lm(sale_price ~ ., data = as.data.frame(st_drop_geometry(regression_data)))

summary(regression)

```
