---
title: "Deliverables_Template"
author: "Dave Drennan"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r deliverables packages}

# install.packages("stargazer") 
# install.packages("spdep")
# install.packages("caret")
# install.packages("FNN")
# install.packages("ggcorrplot")
# install.packages("ckanr")

#https://web.northeastern.edu/econpress/wp-content/uploads/2016/04/Stargazer.pdf
library(stargazer) #remember to include citations for packages

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)

root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")

```

```{r load data}

#REPLACEME

#Confirm clean data

```

```{r train and test sets}

#REPLACEME

```


Deliverable

# Introduction

# Summary Statistics

```{r stargazer table}

#stargazer(REPLACEME, type="text", title="Table 1: Summary Statistics")


```


# Correlation Matrix

Example from book for matrix

```{r correlation matrix}

numericVars <- 
  select_if(st_drop_geometry(training), is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#25CB10", "white", "#FA7800"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation across numeric variables") 

```

# Home Price Scatterplots

Example from book for facetwrap scatterplots

```{r scatterplots}

st_drop_geometry(training) %>% 
  select(sale_price, total_livable_area, age, total_area) %>%
  filter(sale_price <= 1000000, age < 500) %>%
  gather(Variable, Value, -sale_price) %>% 
   ggplot(aes(Value, sale_price)) +
     geom_point(size = .5) + geom_smooth(method = "lm", se=F, colour = "#FA7800") +
     facet_wrap(~Variable, ncol = 3, scales = "free") +
     labs(title = "Price as a function of continuous variables") +
     plotTheme()

```

# Map Home Prices

Follow book - map quintile breaks over neighborhoods 
```{r map home prices}
nhoods <- 
  st_read("data/neighborhoods.geojson") %>%
  st_transform(coordinate_system)

#book used price per square foot
ggplot() +
  geom_sf(data = nhoods, fill = "grey40") +
  geom_sf(data = training, aes(colour = q5(sale_price)), 
          show.legend = "point", size = .75) +
  scale_colour_manual(values = palette5,
                   labels=qBr(training,"sale_price"),
                   name="Quintile\nBreaks") +
  labs(title="Sale Price, Philly") +
  mapTheme()


```

# 3 Interesting Maps (bonus for something extra engaging)

Example of heat map from book

```{r heatmap}

training_age <-
  training %>%
    filter(age < 500) %>%
    na.omit() %>%
    st_as_sf(coords = c("Long", "Lat"), crs = 4326, agr = "constant") %>%
    st_transform(coordinate_system) %>%
    distinct()

ggplot() + geom_sf(data = nhoods, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(training_age)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#25CB10", high = "#FA7800", 
                      breaks=c(0.000000003,0.00000003),
                      labels=c("Minimum","Maximum"), name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Home Age, Philly") +
  mapTheme()
```

# Methods

# Table of Results (Training Set)

Example from book

```{r regression training set}

reg1 <- lm(SalePrice ~ ., data = st_drop_geometry(boston.sf) %>% 
                                 dplyr::select(SalePrice, LivingArea, Style, 
                                               GROSS_AREA, R_TOTAL_RM, NUM_FLOORS,
                                               R_BDRMS, R_FULL_BTH, R_HALF_BTH, 
                                               R_KITCH, R_AC, R_FPLACE))
summary(reg1)

```

# Table of Goodness of Fit (Test Set)

# CV Results

Example from book

```{r cross validation}

fitControl <- trainControl(method = "cv", number = 100)
set.seed(825)

reg.cv <- 
  train(sale_price ~ ., data = st_drop_geometry(training) %>% 
                                dplyr::select(sale_price, 
                                number_stories, 
                                number_of_bathrooms, 
                                number_of_bedrooms), 
     method = "lm", trControl = fitControl, na.action = na.pass)

reg.cv

```

# Pred vs Obs Scatterplot

Example from book

```{r pred and obs}
bothRegressions %>%
  dplyr::select(SalePrice.Predict, SalePrice, Regression) %>%
    ggplot(aes(SalePrice, SalePrice.Predict)) +
  geom_point() +
  stat_smooth(aes(SalePrice, SalePrice), 
             method = "lm", se = FALSE, size = 1, colour="#FA7800") + 
  stat_smooth(aes(SalePrice.Predict, SalePrice), 
              method = "lm", se = FALSE, size = 1, colour="#25CB10") +
  facet_wrap(~Regression) +
  labs(title="Predicted sale price as a function of observed price",
       subtitle="Orange line represents a perfect prediction; Green line represents prediction") +
  plotTheme()

```

# Residuals Map w/ Moran's I

# Predicted Map (Test Set)

# MAPE map by neighborhood

Example from book

```{r mape map}
st_drop_geometry(bothRegressions) %>%
  group_by(Regression, Name) %>%
  summarize(mean.MAPE = mean(SalePrice.APE, na.rm = T)) %>%
  ungroup() %>% 
  left_join(nhoods) %>%
    st_sf() %>%
    ggplot() + 
      geom_sf(aes(fill = mean.MAPE)) +
      geom_sf(data = bothRegressions, colour = "black", size = .5) +
      facet_wrap(~Regression) +
      scale_fill_gradient(low = palette5[1], high = palette5[5],
                          name = "MAPE") +
      labs(title = "Mean test set MAPE by neighborhood") +
      mapTheme()

```

# Scatterplot - MAPE by neighborhood mean price



# Split by Census Groups

Example from book

```{r tracts}
tracts17 <- 
  get_acs(geography = "tract", variables = c("B01001_001E","B01001A_001E","B06011_001"), 
          year = 2017, state=25, county=025, geometry=T, output = "wide") %>%
  st_transform('ESRI:102286')  %>%
  rename(TotalPop = B01001_001E,
         NumberWhites = B01001A_001E,
         Median_Income = B06011_001E) %>%
  mutate(percentWhite = NumberWhites / TotalPop,
         raceContext = ifelse(percentWhite > .5, "Majority White", "Majority Non-White"),
         incomeContext = ifelse(Median_Income > 32322, "High Income", "Low Income"))

grid.arrange(ncol = 2,
  ggplot() + geom_sf(data = na.omit(tracts17), aes(fill = raceContext)) +
    scale_fill_manual(values = c("#25CB10", "#FA7800"), name="Race Context") +
    labs(title = "Race Context") +
    mapTheme() + theme(legend.position="bottom"), 
  ggplot() + geom_sf(data = na.omit(tracts17), aes(fill = incomeContext)) +
    scale_fill_manual(values = c("#25CB10", "#FA7800"), name="Income Context") +
    labs(title = "Income Context") +
    mapTheme() + theme(legend.position="bottom"))

#MAPE for both regressions by race
st_join(bothRegressions, tracts17) %>% 
  group_by(Regression, raceContext) %>%
  summarize(mean.MAPE = scales::percent(mean(SalePrice.APE, na.rm = T))) %>%
  st_drop_geometry() %>%
  spread(raceContext, mean.MAPE) %>%
  kable(caption = "Test set MAPE by neighborhood racial context")

#MAPE for both regressions by income
st_join(bothRegressions, tracts17) %>% 
  filter(!is.na(incomeContext)) %>%
  group_by(Regression, incomeContext) %>%
  summarize(mean.MAPE = scales::percent(mean(SalePrice.APE, na.rm = T))) %>%
  st_drop_geometry() %>%
  spread(incomeContext, mean.MAPE) %>%
  kable(caption = "Test set MAPE by neighborhood income context")

```

# Discussion
## Accuracy

## Generalizability

# Conclusion and Recommendation